---
title: "HW5"
format: 
  html:
    toc: true
    toc-location: left
    code-fold: true
    theme: yeti 
execute:
  message: false
  warning: false
---

Ethan Blacher

6/5/23

Homework 5

```{r Install_Packages}
# should haves (from last week)
library(tidyverse)
library(here)
library(janitor)
library(ggeffects)
library(performance)
library(naniar) # or equivalent
library(flextable) # or equivalent
library(car)
library(broom)
# would be nice to have
library(corrplot)
library(AICcmodavg)
library(GGally)
```

6.  **Introduction**

    This analysis revolves around a study conducted at Harvard Forest in 2005, which explored the impact of prey availability on Sarracenia physiology. Sarracenia, also known as pitcher plants, are carnivorous wetland flora

    This study addressed the effects of prey availability on a litany of factors, such as the photosynthetic rate (Amass), chlorophyll fluorescence, growth, architecture, and foliar nutrient and chlorophyll content of ten pitcher plant (Sarracenia) species. We tested the hypothesis that increased prey availability would stimulate Amass of one or more leaves, increase photosynthetic N- and P-use efficiencies (PNUEN, PNUEP), increase relative biomass allocation to photosynthetically efficient, non-predatory phyllodes rather than pitchers, increase overall plant biomass, and reduce stress-related chlorophyll fluorescence.

    7. **Methods**

    ```{r}

    ```

    \

8.  **Results**

```{r Read_Data}
sarr <- read_csv(here::here("data.hw5","knb-lter-hfr.109.18 (3)", "hf109-01-sarracenia.csv")) %>% 
  #make the column names clear
  clean_names() %>% 
  #selecting the columns of interest
  select(totmass, species, feedlevel, sla, chlorophyll, amass, num_lvs, num_phylls)
```

```{r Missing_Data_Viz}
gg_miss_var(sarr)

```

```{r Drop_NA}
sarr_subset <- sarr %>% 
  drop_na(sla, chlorophyll, amass, num_lvs, num_phylls)
```

```{r Pearson_Viz}
#calculate pearson's r only for numerical values only
sarr_cor <- sarr_subset %>% 
  select(feedlevel:num_phylls) %>% 
  cor(method = "pearson")

#creating a correlation plot
corrplot(sarr_cor,
         #change the shape of what's in the cells
         #point to the left=negative relationship, right=positive relationship
         method = "ellipse",
         #adds the correlation values in black over the elipses
         addCoef.col= "black")

```
